What's in a name? On the (automated) topic naming of software maintenance activities

### 2010-03-03 icpc/icpc2010.tex
minimal edit
generating the data

One issue that arises with use of unsupervised techniques is how to label the topics. While the topic models themselves are generated automatically, what to make of them is less clear. For example, in our previous work~\cite{Hindle09ICSM}, and in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly. E.g., for the word list ``listener change remove add fire", Baldi et al. assign the keyword `event-handling'. The labels are reasonable enough, but still require some expert in the field to determine them. Our technique is automated, because we match keywords from Wordnet to words in the topic model.

@@

\section{Methodology}
 % To gather the data, we first extract the commit log comments from a
% project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
% into 30-day non-overlapping windows. Each window is subject to LDA analysis. %Figure \ref{fig:commits}
% depicts the general process for processing and analyzing the commit messages.

% XXX how many do we have
We wanted to compare two mature software projects, that had open repositories, within the same domain. This allows us to partially control for differences in functional requirements. We selected MySQL and MaxDB, open-source, near-commercial quality database software. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's `SLOCCount'.}. There are approximately ... developers working on it. MySQL 3.23, released in early 2001, contains 320 thousand lines of C and C++ source code. The MySQL project started in 1994.

\subsection{Generating the data}
From each project, we used source code commit data (the message attached to the commit). We leveraged the data we gathered in \cite{Hindle09ICSM} for this work. A typical commit message is: ``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course) ". We extracted these messages and indexed them by creation time. We apply stop-word removal to eliminate common English words like `the', `at', and so on.

From that dataset, we created an XML file which grouped commits into temporal periods, each 30 days wide. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. Given a text file full of commits, indexed by date and time, we then ran an LDA modeler on each period, to generate a list of topics per period. E.g., in Period 1, the most important topic might have been ``security", with ``sql'' second, down on to less relevant topics. We arbitrarily set the number of topics to generate to 20, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics.

\subsection{Associating labels}
% We would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period.
Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms. Recall that topics are lists of words and associated frequencies. We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words. We used several different word lists for comparison.

>>>>

One issue that arises with use of unsupervised techniques is how to label the topics. While the topic models themselves are generated automatically, what to make of them is less clear. For example, in our previous work~\cite{Hindle09ICSM}, and in \cite{Baldi2008}, topics are named manually: human experts read the highest-frequency members of a topic and assign a keyword accordingly. E.g., for the word list ``listener change remove add fire'', Baldi et al. assign the keyword \emph{event-handling}. The labels are reasonable enough, but still require some expert in the field to determine them. Our technique is automated, because we match keywords from Wordnet to words in the topic model.

@@

\section{Methodology}

\begin{comment}
% To gather the data, we first extract the commit log comments from a
% project's SCS repository. We filter out stop words and produce word distributions from these messages. These distributions are bucketed
% into 30-day non-overlapping windows. Each window is subject to LDA analysis. %Figure \ref{fig:commits}
% depicts the general process for processing and analyzing the commit messages.
\end{comment}

% XXX how many do we have
We wanted to compare two mature software projects, that had open repositories, within the same application domain. This allows us to partially control for differences in functional requirements. We selected MySQL and MaxDB, open-source, near-commercial quality database software. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code~\footnote{generated using David A. Wheeler's \emph{SLOCCount}.}.
%How many developers?
%There are approximately ... developers working on it. 
%XXX version #?
The MySQL project started in 1994 but MySQL 3.23 was released in early 2001. MySQL contains 320 thousand lines of C and C++ source code.

\subsection{Generating the data}
From each project, we used source control commit comments, these are the messages that programmers write when they commit revisions to a source control repository. We leveraged the data that we gathered in \cite{Hindle09ICSM} for this work. An example of a typical commit message is: ``history annotate diffs bug fixed (if mysql\_real\_connect() failed there were two pointers to malloc'ed strings, with memory corruption on free(), of course)''. We extracted these messages and indexed them by creation time. We summarized each message as a word distribution but removed stop-words such as   common English words like \emph{the} and \emph{at}.

From that dataset, we created an XML file which windowed commits into
monthly windows. This size of period is smaller than the time between minor releases but large enough for there to be sufficient commits to analyze. We applied an Blei's LDA implementation against the word distributions of these commits, and generated list of topics per period. 
%A topic extracted in the first period Period 1, the most important %topic might have been ``security", with ``sql'' second, down on to %less relevant topics. 
We arbitrarily set the number of topics to generate to $20$, because past experimentation showed that fewer topics might aggregate multiple unique topics while any more topics seemed to dilute the results and create indistinct topics. As well, more than $20$ topics quickly became infeasible for inspection and it was difficult to discern the difference in topics.

\subsection{Associating labels}
\begin{comment}
% We would like to `name' our topic models for each period. In this way, we can see how certain words -- subjects -- occur more or less often with each period.
\end{comment}

Topics are word distributions, essentially ranked lists of words, thus they can be burdensome to interpret and hard to distinguish.
Once we had topics for each period, we tried to associate them with a label from a list of keywords and related terms. Recall that topics are lists of words and associated frequencies. We performed simple string matching between these topics and our lists, `naming' a topic if it contained that word or words. We used several different word lists for comparison.
